Equivalence Hypergraphs: E-Graphs for Monoidal Theories

### 2025-01-23 LICS_2025/monoidal_e_graphs/sections/appendix.tex
wip

\section{$\catname{SLat}$}
\label{sec:appendix:slat}

In this section we will define the category of semilattices that we use as a base for enrichment throughout the paper.

@@

    \begin{proposition}[Special case of Proposition 6.4.6~\cite{Borceux_1994}]
      The forgetful functor $U : \catname{SLat} \to \catname{Set}$ has a left adjoint free functor $F : \catname{Set} \to \catname{SLat}$.
    \end{proposition}
    \begin{proof}
      The functor $F$ is defined by letting $F(A) = \coprod_{A} I_{\catname{SLat}}$ where the latter is a coproduct which is a `free' product of semilattices defined as follows.
      The elements of $S_{1} \coprod S_{2}$ are sequences $s_{1} + s_{2} + \ldots + s_{n}$ where each $s_{i}$ is either from $S_{1}$ or $S_{2}$ quotiented by all relations in $S_{1}$ and $S_{2}$.
      The adjunction is then given by the following natural isomorphisms
      \[
      \catname{SLat}(\coprod_{A}(I_{\catname{SLat}}), B) \cong \prod_{A}(\catname{SLat}(I_{\catname{SLat}}, B)) \cong \catname{Set}(A,\catname{SLat}(I_{\catname{SLat}}, B)) \cong \catname{Set}(A, U(B))
      \]
      The fist isomorphism is given by the fact that hom-functor makes limits into colimits in its first argument, the second being given by the property that $|\catname{Set}(A,B)| = |B|^{|A|} = |\underbrace{B \times \ldots \times B}_{|A|}|$, and the last isomorphism is given by the definition of $U$.
      Furthermore, we have, 
      \[
      F(I) \cong I_{\catname{SLat}}
      \]
      and 
      \begin{align*}
      F(A) \otimes F(B) &\cong (\coprod_{A} I_{\catname{SLat}}) \otimes (\coprod_{B} I_{\catname{SLat}})\\
            &\cong \coprod_{A} (I_{\catname{SLat}} \otimes \coprod_{B} I_{\catname{SLat}})\\
            &\cong \coprod_{A} (\coprod_{B} (I_{\catname{SLat}} \otimes I_{\catname{SLat}}))\\
            &\cong \coprod_{A} (\coprod_{B} I_{\catname{SLat}})\\
            &\cong \coprod_{A \times B} I_{\catname{SLat}}\\
            &\cong F(A \times B)
      \end{align*}
    In the above we used the fact that functors $- \otimes X$ and $X \otimes -$ preserve colimits as they are both left-adjoint by symmetry and monoidal closedness of $\catname{SLat}$.
    \end{proof}

\section{Proofs for Section \ref{sec:e-hypergraphs}: E-hypergraphs}
\label{sec:appendix:pushout}

In this section we will define the conditions under which the pushout in $\catname{EHyp}(\Sigma)$ exists, and we will explicitly construct such a pushout.
We will also show the uniqueness of a boundary pushout complement.
These condition will play a crucial role in showing that for each enrichment equation in $\textbf{PROP}^{+}(\Sigma)$ there is a sequence of rewrites in $\WellTypedMdaEcospans$.

@@

We then quotient the set vertices and edges in $X + Y$ by $R_{V}$ and $R_{E}$ respectively.
\ifdefined \ONECOLUMN
\begin{align*}
    \sfrac{X + Y}{\sim (R_{V},R_{E})} = \{
        &\sfrac{V_{X} + V_{Y}}{\sim (R_{V},R_{E})}, \sfrac{E_{X} + E_{Y}}{\sim (R_{V},R_{E})}, \sfrac{s_{X+Y}}{\sim (R_{V},R_{E})},\\
        &\sfrac{t_{X,Y}}{\sim (R_{V},R_{E})}, \consistency_{\sfrac{X+Y}{\sim (R_{V},R_{E})}}, <_{\sfrac{X + Y}{\sim (R_{V},R_{E})}}\}    
\end{align*}
\else
    \begin{align*}
        \sfrac{X + Y}{\sim (R_{V},R_{E})} = \{&\\
            &\sfrac{V_{X} + V_{Y}}{\sim (R_{V},R_{E})}, \sfrac{E_{X} + E_{Y}}{\sim (R_{V},R_{E})}, \sfrac{s_{X+Y}}{\sim (R_{V},R_{E})},\\
            &\sfrac{t_{X,Y}}{\sim (R_{V},R_{E})}, \consistency_{\sfrac{X+Y}{\sim (R_{V},R_{E})}}, <_{\sfrac{X + Y}{\sim (R_{V},R_{E})}}\\
        & \}    
    \end{align*}
\fi
We will refer to $\sim (R_{V},R_{E})$ just as $\sim$, \textit{e.g.}, by writing $\sfrac{V_{X} + V_{Y}}{\sim}$ and concrete relation will be clear from the context.
Where necessary we will refer to $S_{V}$ and $S_{R}$ as $\sim_{S}$.
We have 
\[
    \sfrac{s_{X+Y}}{\sim} : \sfrac{E_{X} + E_{Y}}{\sim} \to (\sfrac{V_{X} + V_{Y}}{\sim})^{*}
\]
and there is an obvious surjective function $[]_{V} : (V_{X} + V_{Y}) \to (\sfrac{V_{X} + V_{Y}}{\sim})$ that maps elements to their equivalence classes and $[]_{V}^{*}$ is its extension to sequences.
There is also $[] : E_{X} + E_{Y} \to \sfrac{E_{X} + E_{Y}}{\sim}$ and we will omit subscripts as the correct type will be clear from the argument.
We then define 
\[
    \sfrac{s_{X+Y}}{\sim}([e]) = s_{X+Y};[]^{*}(e) = [s_{X+Y}(e)]^{*}
\]
We will also use subscripts when it is important to tell if an element of $E_{X} + E_{Y}$ has a pre-image in either $E_{X}$ or $E_{Y}$ by writing $e_{x}$ or $e_{y}$.
Similarly, we will write $v_{x}$ to refer to a vertex with a pre-image in $V_{X}$.
Likewise, 
\[
    t_{\sfrac{X+Y}{\sim}}([e]) = [t_{X+Y}(e)]^{*}
\]

@@

                u_{V}^{*}([s_{X+Y}(e)]^{*}) = u_{V}^{*}([s_{X+Y}(\iota_{1,E}(e'))]^{*}) = u_{V}^{*}([(\iota_{1,V}^{*}(s_{X}(e')))]^{*})

@@

    [u_{V};\iota_{V_{Q}}, u_{E};\iota_{E_{Q}}]^{*}(\consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))) \subseteq \consistency_{Q}(u_{V};\iota_{V_{Q}}([v]))

@@

      [u_{V};\iota_{V_{Q}}, u_{E};\iota_{E_{Q}}]^{*}(\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))) \subseteq \consistency_{Q}(u_{V};\iota_{V_{Q}}([v]))

@@

            [u_{V};\iota_{Q_{V}},u_{E};\iota_{Q_{E}}]^{*}(\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([u]))) \subseteq \consistency_{Q}(\iota_{V_{Q}}(u_{V}([u])))

@@

        [u_{V};\iota_{V_{Q}}, u_{E};\iota_{E_{Q}}]^{*}(\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))) \subseteq \consistency_{Q}(u_{V};\iota_{V_{Q}}([v]))

@@

        [u_{V};\iota_{V_{Q}}, u_{E};\iota_{E_{Q}}]^{*}(\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e]))) \subseteq \consistency_{Q}(u_{E};\iota_{E_{Q}}([e]))
       \] follows by the same argument.

@@

        [u_{V};\iota_{V_{Q}}, u_{E};\iota_{E_{Q}}]^{*}(\consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))) \subseteq \consistency_{Q}(u_{V};\iota_{V_{Q}}([v]))

@@

        or 
        \[
            \iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([x_{i+1}]) \in \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([x_{i}]))
        \]
        for $i < n$ (or similarly for edges), then $u_{V};\iota_{V_{Q}}([x_1]) \in \consistency_{Q}(u_{V};\iota_{V_{Q}}([x_n]))$.
        \begin{itemize}
            \item Base case is when $w = (x_1,x_2)$.
                  Then the statement is trivial as 
                  \begin{align*}
                  &[u_{V};\iota_{V_{Q}},u_{E};\iota_{E_{Q}}]^{*}(\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([x_2])))\\
                  &\subseteq\\
                  &\consistency_{Q}(u_{V};\iota_{V_{Q}}([x_2]))
                  \end{align*}
            \item Suppose the statement is true for all $w$ such that $|w| \leq n$.
            \item Consider a sequence $w = (x_1, \ldots, x_{n}, x_{n+1})$.
                  By assumption, we know that $u_{V};\iota_{V_{Q}}([x_1]) \in \consistency_{Q}(u_{V};\iota_{V_{Q}}([x_n]))$.
                  The fact that $\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([x_{n}]) \in \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([x_{n+1}]))$ (or its symmetric counterpart) implies that $u_{V};\iota_{V_{Q}}([x_{n}]) \in \consistency_{Q}(u_{V};\iota_{V_{Q}}([x_{n+1}]))$ and by the transitivity and symmetry of $\consistency_{Q}$ it follows that
                \[
                    u_{V};\iota_{V_{Q}}([x_1]) \in \consistency_{Q}(u_{V};\iota_{V_{Q}}([x_{n+1}]))
                \]
                and
                \[
                    u_{V};\iota_{V_{Q}}([x_{n+1}]) \in \consistency_{Q}(u_{V};\iota_{V_{Q}}([x_{1}]))
                \]
        \end{itemize}
       \end{itemize}
       Therefore, we have
       \[
        [u_{V};\iota_{V_{Q}}, u_{E};\iota_{E_{Q}}]^{*}(\consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))) \subseteq \consistency_{Q}(u_{V};\iota_{V_{Q}}([v]))
       \]
       and similarly for edges. Which concludes that $(u_{V},u_{E})$ is a homomorphism.
    \end{enumerate}
\end{proof}

@@

                                    <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1])) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v''])) = [e_2]

@@

                <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1])) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v''])) = [e_2]

@@

@@

                        \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) &= [[]_{V}^{\consistency},[]_{E}^{\consistency}]^{*}(\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v'')))\\

@@

                              \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[v_1']) &= [[]_{V}^{\hashtag},[]_{E}^{\hashtag}]^{*}(\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v_1')))\\
                                                                                                                     &= [[]_{V}^{\hashtag},[]_{E}^{\hashtag}]^{*}(\consistency_{X+Y}(\iota_{E_{X} + E_{Y}}(e_1)))\\
                                                                                                                     &= \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1]))\\
                                                                                                                     &= \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[v_1])

@@

                                \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[v_1']) = [[]_{V}^{\hashtag},[]_{E}^{\hashtag}]^{*}(\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v_1)))

@@

                                \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}[e_1]) = [[]_{V}^{\hashtag},[]_{E}^{\hashtag}]^{*}(\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v_1)))

@@

@@

    \textcolor{red}{The same reasoning applies to show that $E_{\mathcal{L}^{\bot}}$} is unique and hence the sets of edges and vertices are uniquely determined in the pushout complement.

@@

    which implies $<_{2}(\iota_{V_{\mathcal{L}^{\bot}}}(v)) = <_{1}(\iota_{V_{\mathcal{L}^{\bot}}}(v))$ for all $v$ such that $[v) \not = \varnothing$ since $g_{E}$ \textcolor{red}{only identifies edges with no successors}.

@@

\section{Proofs for Section \ref{sec:soundness-and-completeness}: Soundness and Completeness}

In this section we will elaborate on soundness of our translation $\llbracket - \rrbracket : \textbf{PROP}^{+}(\Sigma) \to \WellTypedMdaEcospans / \mathcal{S}$.

>>>>

\subsection{$\catname{SLat}$}
\label{sec:appendix:slat}

In this section we define the category of semilattices that we use as a base for enrichment throughout the paper.

@@

    \begin{proposition}[Special case of Proposition 6.4.6~\cite{Borceux_1994}]
      The forgetful functor $U : \catname{SLat} \to \catname{Set}$ has a left adjoint free functor $F : \catname{Set} \to \catname{SLat}$.
    \end{proposition}
    \begin{proof}
      The functor $F$ is defined by letting $F(A) = \coprod_{A} I_{\catname{SLat}}$ where the latter is a coproduct which is a `free' product of semilattices defined as follows.
      The elements of $S_{1} \coprod S_{2}$ are sequences $s_{1} + s_{2} + \ldots + s_{n}$ where each $s_{i}$ is either from $S_{1}$ or $S_{2}$ quotiented by all relations in $S_{1}$ and $S_{2}$.
      The adjunction is then given by the following natural isomorphisms
      \begin{align*}
      \catname{SLat}(\coprod_{A}(I_{\catname{SLat}}), B) &\cong \prod_{A}(\catname{SLat}(I_{\catname{SLat}}, B))\\
                                                         &\cong \catname{Set}(A,\catname{SLat}(I_{\catname{SLat}}, B))\\
                                                         &\cong \catname{Set}(A, U(B))
      \end{align*}
      The fist isomorphism is given by the fact that hom-functor makes limits into colimits in its first argument, the second being given by the property that $|\catname{Set}(A,B)| = |B|^{|A|} = |\underbrace{B \times \ldots \times B}_{|A|}|$, and the last isomorphism is given by the definition of $U$.
      Furthermore, we have, 
      \[
      F(I) \cong I_{\catname{SLat}}
      \]
      and 
      \begin{align*}
      F(A) \otimes F(B) &\cong (\coprod_{A} I_{\catname{SLat}}) \otimes (\coprod_{B} I_{\catname{SLat}})\\
            &\cong \coprod_{A} (I_{\catname{SLat}} \otimes \coprod_{B} I_{\catname{SLat}})\\
            &\cong \coprod_{A} (\coprod_{B} (I_{\catname{SLat}} \otimes I_{\catname{SLat}}))\\
            &\cong \coprod_{A} (\coprod_{B} I_{\catname{SLat}})\\
            &\cong \coprod_{A \times B} I_{\catname{SLat}}\\
            &\cong F(A \times B)
      \end{align*}
    In the above we used the fact that functors $- \otimes X$ and $X \otimes -$ preserve colimits as they are both left-adjoint by symmetry and monoidal closedness of $\catname{SLat}$.
    \end{proof}

    \begin{definition}
        Given two $\catname{SLat}$-functors $F,G : \mathbb{C} \to \mathbb{D}$ a $\catname{SLat}$-natural transformation is family of morphisms indexed by objects in $\mathbb{C}$, $\alpha_{A} : I_{\catname{SLat}} \to \mathbb{D}(FA,GA)$ in $\catname{SLat}$ such that the following diagram commutes
        \[
        \adjustbox{width=\linewidth}{
        \begin{tikzcd}
          & {\mathbb{C}(A,A')\otimes I_{\mathcal{V}}} && {\mathbb{D}(FA,FA') \otimes \mathbb{D}(FA',GA')} \\
          {\mathbb{C}(A,A')} &&&& {\mathbb{D}(FA,GA')} \\
          & {I_{\mathcal{V}} \otimes \mathbb{C}(A,A')} && {\mathbb{D}(FA,GA) \otimes \mathbb{D}(GA,GA')}
          \arrow["{G_{A,A'}\otimes \alpha_{A'}}"', from=1-2, to=1-4]
          \arrow["{c_{FA,FA',GA'}}"', from=1-4, to=2-5]
          \arrow["{r^{-1}}", from=2-1, to=1-2]
          \arrow["{l^{-1}}", from=2-1, to=3-2]
          \arrow["{\alpha_{A} \otimes F_{A,A'}}", from=3-2, to=3-4]
          \arrow["{c_{FA,GA,GA'}}", from=3-4, to=2-5]
      \end{tikzcd}}
      \]
      \end{definition}

      Intuitively, in the case of free enrichment $\alpha_{A}$ picks a usual natural transformation from $\mathbb{D}(FA,GA)$.

      \begin{example}
        Consider a copy natural transformation $\triangle_{A} : A \to A \times A$ of $\catname{S}(\Sigma_{C},\mathcal{E}_{C})$.
        The later Cartesian PROP is lifted to a Cartesian $\catname{SLat}$-PROP where the naturality condition can be traced component-wise where ${*}$ is a one element semilattice:

      \[\adjustbox{width=\linewidth}{
      \begin{tikzcd}  
          & {f + g \otimes *} && {\triangle_{A} \otimes ((f+g) \times (f+g))} \\
          f + g &&&& {\triangle_{A'};((f + g) \times (f+g)) = (f + g);\triangle_{A'};} \\
          & {* \otimes f + g} && {f + g \otimes \triangle_{A'}}
          \arrow[from=1-2, to=1-4]
          \arrow[from=1-4, to=2-5]
          \arrow[from=2-1, to=1-2]
          \arrow[from=2-1, to=3-2]
          \arrow[from=3-2, to=3-4]
          \arrow[from=3-4, to=2-5]
      \end{tikzcd}}
      \]

      which is the usual naturality condition.

      \end{example}

\subsection{Proofs for Section \ref{sec:e-hypergraphs}: E-hypergraphs}
\label{sec:appendix:pushout}

In this section we define the conditions under which the pushout in $\catname{EHyp}(\Sigma)$ exists, and we explicitly construct such a pushout.
We also show the uniqueness of a boundary pushout complement.
These condition will play a crucial role in showing that for each enrichment equation in $\textbf{PROP}^{+}(\Sigma)$ there is a sequence of rewrites in $\WellTypedMdaEcospans$.

@@

We then quotient the set vertices and edges in $X + Y$ by $R_{V}$ and $R_{E}$ respectively.
\ifdefined \ONECOLUMN
\begin{align*}
    \sfrac{X + Y}{\sim (R_{V},R_{E})} = \{
        &\sfrac{V_{X} + V_{Y}}{\sim (R_{V},R_{E})}, \sfrac{E_{X} + E_{Y}}{\sim (R_{V},R_{E})}, \sfrac{s_{X+Y}}{\sim (R_{V},R_{E})},\\
        &\sfrac{t_{X,Y}}{\sim (R_{V},R_{E})}, \consistency_{\sfrac{X+Y}{\sim (R_{V},R_{E})}}, <_{\sfrac{X + Y}{\sim (R_{V},R_{E})}}\}    
\end{align*}
\else
    \begin{align*}
        \sfrac{X + Y}{\sim (R_{V},R_{E})} = (&\\
            &\sfrac{V_{X} + V_{Y}}{\sim (R_{V},R_{E})},\\
            &\sfrac{E_{X} + E_{Y}}{\sim (R_{V},R_{E})},\\
            &\sfrac{s_{X+Y}}{\sim (R_{V},R_{E})},\\
            &\sfrac{t_{X,Y}}{\sim (R_{V},R_{E})},\\
            &\consistency_{\sfrac{X+Y}{\sim (R_{V},R_{E})}},\\
            &<_{\sfrac{X + Y}{\sim (R_{V},R_{E})}}\\
        &)    
    \end{align*}
\fi
We will refer to $\sim (R_{V},R_{E})$ just as $\sim$, \textit{e.g.}, by writing $\sfrac{V_{X} + V_{Y}}{\sim}$ and concrete relation will be clear from the context.
Where necessary we will refer to $S_{V}$ and $S_{R}$ as $\sim_{S}$.
We have 
\[
    \sfrac{s_{X+Y}}{\sim} : \sfrac{E_{X} + E_{Y}}{\sim} \to (\sfrac{V_{X} + V_{Y}}{\sim})^{*}
\]
and there is an obvious surjective function $[]_{V} : (V_{X} + V_{Y}) \to (\sfrac{V_{X} + V_{Y}}{\sim})$ that maps elements to their equivalence classes and $[]_{V}^{*}$ is its extension to sequences.
There is also $[] : E_{X} + E_{Y} \to \sfrac{E_{X} + E_{Y}}{\sim}$ and we will omit subscripts as the correct type will be clear from the argument.
We then define 
\[
    \sfrac{s_{X+Y}}{\sim}([e]) = s_{X+Y};[]^{*}(e) = [s_{X+Y}(e)]^{*}
\]
We will also use subscripts when it is important to tell if an element of $E_{X} + E_{Y}$ has a pre-image in either $E_{X}$ or $E_{Y}$ by writing $e_{x}$ or $e_{y}$.
Similarly, we will write $v_{x}$ to refer to a vertex with a pre-image in $V_{X}$.
Likewise, 
\[
    t_{\sfrac{X+Y}{\sim}}([e]) = [t_{X+Y}(e)]^{*}
\]

@@

                u_{V}^{*}([s_{X+Y}(e)]^{*}) &= u_{V}^{*}([s_{X+Y}(\iota_{1,E}(e'))]^{*})\\
                                            &= u_{V}^{*}([(\iota_{1,V}^{*}(s_{X}(e')))]^{*})

@@

    &[u_{V};\iota_{V_{Q}}, u_{E};\iota_{E_{Q}}]^{*}(\consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])))\\
    &\;\subseteq\\
    &\consistency_{Q}(u_{V};\iota_{V_{Q}}([v]))

@@

      &\;\subseteq\\
      &\consistency_{Q}(u_{V};\iota_{V_{Q}}([v]))

@@

            &[u_{V};\iota_{Q_{V}},u_{E};\iota_{Q_{E}}]^{*}(\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([u])))\\
            &\;\subseteq\\
            &\consistency_{Q}(\iota_{V_{Q}}(u_{V}([u])))

@@

        &\;\subseteq\\
        & \consistency_{Q}(u_{V};\iota_{V_{Q}}([v]))

@@

        &[u_{V};\iota_{V_{Q}}, u_{E};\iota_{E_{Q}}]^{*}(\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])))\\
        &\;\subseteq\\
        &\consistency_{Q}(u_{E};\iota_{E_{Q}}([e]))
       \end{align*} follows by the same argument.

@@

        &[u_{V};\iota_{V_{Q}}, u_{E};\iota_{E_{Q}}]^{*}(\consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])))\\
        &\; \subseteq\\
        &\consistency_{Q}(u_{V};\iota_{V_{Q}}([v]))

@@

        or 
        \[
            \iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([x_{i+1}]) \in \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([x_{i}]))
        \]
        for $i < n$ (or similarly for edges), then $u_{V};\iota_{V_{Q}}([x_1]) \in \consistency_{Q}(u_{V};\iota_{V_{Q}}([x_n]))$.
        \begin{itemize}
            \item Base case is when $w = (x_1,x_2)$.
                  Then the statement is trivial as 
                  \begin{align*}
                  &[u_{V};\iota_{V_{Q}},u_{E};\iota_{E_{Q}}]^{*}(\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([x_2])))\\
                  &\subseteq\\
                  &\consistency_{Q}(u_{V};\iota_{V_{Q}}([x_2]))
                  \end{align*}
            \item Suppose the statement is true for all $w$ such that $|w| \leq n$.
            \item Consider a sequence $w = (x_1, \ldots, x_{n}, x_{n+1})$.
                  By assumption, we know that $u_{V};\iota_{V_{Q}}([x_1]) \in \consistency_{Q}(u_{V};\iota_{V_{Q}}([x_n]))$.
                  The fact that $\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([x_{n}]) \in \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([x_{n+1}]))$ (or its symmetric counterpart) implies that $u_{V};\iota_{V_{Q}}([x_{n}]) \in \consistency_{Q}(u_{V};\iota_{V_{Q}}([x_{n+1}]))$ and by the transitivity and symmetry of $\consistency_{Q}$ it follows that
                \[
                    u_{V};\iota_{V_{Q}}([x_1]) \in \consistency_{Q}(u_{V};\iota_{V_{Q}}([x_{n+1}]))
                \]
                and
                \[
                    u_{V};\iota_{V_{Q}}([x_{n+1}]) \in \consistency_{Q}(u_{V};\iota_{V_{Q}}([x_{1}]))
                \]
        \end{itemize}
       \end{itemize}
       Therefore, we have
       \begin{align*}
        &[u_{V};\iota_{V_{Q}}, u_{E};\iota_{E_{Q}}]^{*}(\consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])))\\
        &\;\subseteq\\
        &\consistency_{Q}(u_{V};\iota_{V_{Q}}([v]))
       \end{align*}
       and similarly for edges. Which concludes that $(u_{V},u_{E})$ is a homomorphism.
    \end{enumerate}
\end{proof}

@@

                                    <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1])) &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v'']))\\

@@

                <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1])) &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v'']))\\
                 &\;= [e_2]

@@

                        \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1])) &=\\
                        \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1]))&

@@

                        \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}&(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) =\\
                                                                                                                &= [[]_{V}^{\consistency},[]_{E}^{\consistency}]^{*}(\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v'')))\\

@@

                              \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}&(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[v_1']) =\\
                              &\;= [[]_{V}^{\hashtag},[]_{E}^{\hashtag}]^{*}(\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v_1')))\\
                                                                                                                     &\;= [[]_{V}^{\hashtag},[]_{E}^{\hashtag}]^{*}(\consistency_{X+Y}(\iota_{E_{X} + E_{Y}}(e_1)))\\
                                                                                                                     &\;= \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1]))\\
                                                                                                                     &\;= \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[v_1])

@@

                                &\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[v_1']) =\\
                                &\;[[]_{V}^{\hashtag},[]_{E}^{\hashtag}]^{*}(\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v_1)))

@@

                                &\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}[e_1]) =\\
                                &\;[[]_{V}^{\hashtag},[]_{E}^{\hashtag}]^{*}(\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v_1)))

@@

                                \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}&(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1])) =\\
                                &\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1'']))\\ 

@@

    The same reasoning applies to show that $E_{\mathcal{L}^{\bot}}$ is unique and hence the sets of edges and vertices are uniquely determined in the pushout complement.

@@

    which implies $<_{2}(\iota_{V_{\mathcal{L}^{\bot}}}(v)) = <_{1}(\iota_{V_{\mathcal{L}^{\bot}}}(v))$ for all $v$ such that $[v) \not = \varnothing$ since $g_{E}$ only identifies edges with no successors.

@@

\subsection{Proofs for Section \ref{sec:soundness-and-completeness}: Soundness and Completeness}
\label{sec:appendix:detailed_proofs}
In this section we will elaborate on soundness and completeness of our translation $\llbracket - \rrbracket : \catname{S}^{+}(\Sigma, \mathcal{E}) \to \WellTypedMdaEcospans / \mathcal{S}, \mathcal{E}$.
First we recite the interpretation function $[]$ on $\Sigma$-terms as defined in~\cite{bonchi_string_2022-2} that induces a functor $\catname{S}(\Sigma) \to \MdaCospans$ and eventually a functor $\catname{S}^{+}(\Sigma) \to \MdaCospans^{+}$.
The base cases for interpretation function $[-] : \textbf{SMT}(\Sigma) \to \MdaCospans$ is given in the Figure~\ref{fig:base_cases}.
Then it is defined inductively by letting $[f \otimes g] \Coloneqq [f] \otimes [g]$ and $[f;g] \Coloneqq [f];[g]$ and by freeness induces the corresponding functor.

\begin{remark}\label{remark:embedding_functor}
    There is a faithful identity-on-objects functor $E: \HypI{\Sigma} \to \Ecospans$ which maps a cospan $n \xrightarrow{f} \mathcal{G} \xleftarrow{g} m$ to an extended cospan $n \xrightarrow{f_{ext}} n \xrightarrow{f_{int}} \mathcal{G} \xleftarrow{g_{int}} m \xleftarrow{g_{ext}} m$ such that $f_{ext};f_{int} = f$ and $g_{ext};g_{int} = g$.
\end{remark}

\begin{lemma}
\label{lemma:normal_form}
For each cospan $f$ in ${\WellTypedMdaEcospans}/{\mathcal{S}}$ there is a \textit{normal form} such that 
\[
	f = f_1 + \ldots + f_n
\]
	such that each $f_i$ contains no hierarchical edges,  and for all $i \neq j$ we have $f_i \neq f_j$.
\end{lemma}

\begin{figure*}
\[
\adjustbox{width=\textwidth}{
\tikzfig{../figures/appendix/interpretation}
}
\]
\caption{Base cases for $[-] : \textbf{SMT}(\Sigma) \to \MdaCospans$}
\label{fig:base_cases}
\end{figure*}

\begin{lemma}[Lemma~\ref{lemma:cospans_plus_equiv} restatement]
